---
title: "JKO"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(MASS)
library(ggfortify)
library(cluster)
```

```{r}
library(transport)
library(MEPDF)
library(fishmethods)
library(FNN)
```

```{r}
library(tidyverse)
```

Simple case

Generate data
```{r}
set.seed(525)
n <- 100
x <- sample(x = c(0,1), size = n, replace = TRUE)
x_mx <- matrix(c(rep(1, n), x), ncol = n, byrow = TRUE)
# Get the true beta
mean <- c(0, 0)                                   
cov_mx <- matrix(c(2, 0, 0, 2),
                                   ncol = 2)
  
# create bivariate normal distribution
beta_true <- mvrnorm(n = n,
                        mu = mean, 
                        Sigma = cov_mx)

# beta_0 <- rnorm(1, 0, 2) # -0.6
# beta_1 <- rnorm(1, 0, 2) # 0.95

logit_inv <- function(x){
  1/(1+exp(-x))
}


pr <- logit_inv(diag(beta_true %*% x_mx))
y <- rbinom(n,1,pr)
y_hat <- 2*y-1

```

Get the Prior
```{r}
prior_mean <- c(0, 0)                                   
prior_cov_mx <- matrix(c(100, 0, 0, 100),
                                   ncol = 2)
  
# create bivariate normal distribution
sample_prior <- mvrnorm(n = 1,
                        mu = prior_mean, 
                        Sigma = prior_cov_mx)

```

Log likelihood
```{r}
likelihood <- function(beta){
  sum(log(1+exp(-y_hat*t(beta)%*%x_mx))) + 1/2*t(beta) %*% solve(prior_cov_mx) %*% beta
}

grad_likelihood <- function(beta){
  Z <- (logit_inv(y_hat*t(beta) %*% x_mx)-1) * y_hat 
  Z %*% t(x_mx) + t(beta) %*% solve(prior_cov_mx) 
}

hessian_likelihood <- function(beta){
  Z <- logit_inv(y_hat*t(beta) %*% x_mx)*(1 - logit_inv(y_hat*t(beta) %*% x_mx))
  x_mx %*% diag(as.vector(Z)) %*% t(x_mx) + solve(prior_cov_mx) 
}
```


$$
F(\rho) = D_{KL}(\rho \| P(\beta)) - E_{\rho}[\log P(D|\beta)]
$$

Let $\rho_0 \sim N(\mu,\Sigma)$
```{r}
rho_0_mean <- c(0, 0)                                   
rho_0_cov_mx <- matrix(c(1000, 0, 0, 1000),
                       ncol = 2)

rho_0  <- mvrnorm(n = 1,
                  mu = rho_0_mean,
                  Sigma = rho_0_cov_mx)

```

First term
```{r}
set.seed(525)

num_sample <- 1000
sample_prior <- mvrnorm(n = num_sample,
                        mu = prior_mean, 
                        Sigma = prior_cov_mx)

rho_0  <- mvrnorm(n = num_sample,
                  mu = rho_0_mean,
                  Sigma = rho_0_cov_mx)

mean(KL.divergence(sample_prior, rho_0))
```

second term
```{r}
apply(rho_0,1,likelihood) %>% mean()
```

get $F(\rho)$,

```{r}
get_F <- function(rho, num_sample = 1000, seed = 525){
  
  set.seed(seed)
  
  sample_prior <- mvrnorm(n = num_sample,
                          mu = prior_mean, 
                          Sigma = prior_cov_mx)
  
  
  first_term <- mean(KL.divergence(sample_prior, rho))
  second_term <- apply(rho,1,likelihood) %>% mean()
  F_val <- first_term - second_term
  F_val
}

```

```{r}
set.seed(525)

rho_0_mean <- c(0, 0)                                   
rho_0_cov_mx <- matrix(c(1000, 0, 0, 1000),
                       ncol = 2)

rho_0  <- mvrnorm(n = num_sample,
                  mu = rho_0_mean,
                  Sigma = rho_0_cov_mx)

get_F(rho_0)
```

Generate samples from empirical distribution
```{r}
plot(density(rho_0[,1]))
plot(density(rho_0[,2]))
```

Method 1
```{r}
get_emp_pdf <- function(x.in, density){
  x.list <- density$x
  y.list <- density$y

  index_list <- c(1:length(x.list))

  index.in <- max(index_list[x.in > x.list ])
  y.in <- y.list[index.in]
  y.in
}

M <- max(density( rho_0[,1])$y )
x_min <- min(density( rho_0[,1])$x )
x_max <- max(density( rho_0[,1])$x )

df2 <- tibble(y = runif(5e2, x_min, x_max), u = runif(5e2, 0, 1),
             loglik_ratio = map_dbl(y, ~get_emp_pdf(.x, density(rho_0[,1]) )) / M,
             selected = u < loglik_ratio)
```

```{r}
f <- function(x) dnorm(x, mean = 0, sd = sqrt(1000))

ggplot(tibble(x = seq(0, 1, 1e-2)), aes(x = x)) +
  stat_function(fun = f) +
  stat_function(fun = function(x) 1/M * dunif(x)) +
  geom_point(aes(x = y, y = M * u, color = selected), data = df2) +
  ylab("Density") + ylim(c(0, 0.013))
```

```{r}
df_selected <- df2 %>%
filter(selected)
# Proportion of selected samples?
NROW(df_selected) / NROW(df2)
NROW(df_selected)
```

```{r}
get_emp_pdf <- function(x.in, density){
  x.list <- density$x
  y.list <- density$y

  index_list <- c(1:length(x.list))

  index.in <- max(index_list[x.in > x.list ])
  y.in <- y.list[index.in]
  y.in
}
```

```{r}
density <- density( rho_0[,1])
total_generated <- 5000
sample_needed <- 1000
  
M <- max(density$y )
x_min <- min(density$x )
x_max <- max(density$x )

df2 <- tibble(y = runif(total_generated, x_min, x_max), u = runif(total_generated, 0, 1),
             loglik_ratio = map_dbl(y, ~get_emp_pdf(.x, density )) / M,
             selected = u < loglik_ratio)

df2_selected <- df2 %>%
  filter(selected)

selected_samples <- df2_selected$y
output <- sample(selected_samples, sample_needed)
plot(density(output))
```
```{r}
generate_samples <- function(density, 
                             total_generated = 6000,
                             sample_needed = 1000,
                             seed = 525){
  set.seed(seed)
  M <- max(density$y )
  x_min <- min(density$x )
  x_max <- max(density$x )
  
  df2 <- tibble(y = runif(total_generated, x_min, x_max), u = runif(total_generated, 0, 1),
               loglik_ratio = map_dbl(y, ~get_emp_pdf(.x, density )) / M,
               selected = u < loglik_ratio)
  
  df2_selected <- df2 %>%
    filter(selected)
  
  selected_samples <- df2_selected$y
  output <- sample(selected_samples, sample_needed)
  output
}
```

```{r}
samples <- generate_samples(density = density( rho_0[,1]))
plot(density(samples))
```

zeroth order gradient estimator
using simultaneous perturbation (Spall (1992))
To estimate $\nabla f(x)$, a sample output would involve first simulating a random vector, say $h = \{ h_1,...,h_p\} \in \{ -1, 1 \}$, then generating two unbiased simulation runs $\hat f (x + \delta h)$ and $\hat f(x - \delta h )$, and finally outputting, for each direction i,
$$
\frac{\hat f (x + \delta h) - \hat f (x - \delta h)}{2 \delta h_i},
$$
where $\delta > 0$ is the perturbation size.


Let 
$$
Loss(\rho) = F(\rho) + \frac{1}{2 \tau} W_2^2 (\rho, \rho^\tau_t).
$$
And the $\rho$ that minimize $Loss(\rho)$ is the $\rho^\tau_{t+1}$.

Loss function

```{r}
set.seed(525)

rho_0_mean <- c(0, 0)                                   
rho_0_cov_mx <- matrix(c(1000, 0, 0, 1000),
                       ncol = 2)

rho_0  <- mvrnorm(n = num_sample,
                  mu = rho_0_mean,
                  Sigma = rho_0_cov_mx)
```


```{r}
rho_finite_element_dim1 <- density(rho_0[,1])
rho_finite_element_dim2 <- density(rho_0[,2])
tau <- 1/60
rho_prev <- rho_0

sample_rho_d1 <- generate_samples(density = rho_finite_element_dim1)
sample_rho_d2 <- generate_samples(density = rho_finite_element_dim2)
sample_rho <- cbind(sample_rho_d1, sample_rho_d2)


# Get F value
F_val <- get_F(sample_rho)

# Get W^2_2
rho_current <- pp(sample_rho)
rho_prev <- pp(rho_prev)
W2 <- wasserstein(rho_current,rho_prev,p=2)^2

loss <- F_val + 1/(2*tau)*W2
```

```{r}
get_loss <- function(rho_finite_element_dim1,
                     rho_finite_element_dim2,
                     rho_prev, tau = 1/60){
  
  sample_rho_d1 <- generate_samples(density = rho_finite_element_dim1)
  sample_rho_d2 <- generate_samples(density = rho_finite_element_dim2)
  sample_rho <- cbind(sample_rho_d1, sample_rho_d2)
  
  
  # Get F value
  F_val <- get_F(sample_rho)
  
  # Get W^2_2
  rho_current <- pp(sample_rho)
  rho_prev <- pp(rho_prev)
  W2 <- wasserstein(rho_current,rho_prev,p=2)^2
  
  loss <- F_val + 1/(2*tau)*W2
  loss
}

```

```{r}
rho_finite_element_dim1 <- density(rho_0[,1])
rho_finite_element_dim2 <- density(rho_0[,2])

get_loss(rho_finite_element_dim1,
         rho_finite_element_dim2,
         rho_prev = rho_0)
```

```{r}

```


```{r}
set.seed(525)

rho_finite_element_dim1 <- density(rho_0[,1])
rho_finite_element_dim2 <- density(rho_0[,2])

delta_dim1 <- max(rho_finite_element_dim1$y)/10
delta_dim2 <- max(rho_finite_element_dim2$y)/10

B <- 100

Loss_gradient_dim1_mx <- matrix(0, ncol = B,
                                nrow = length(rho_finite_element_dim1$y))
Loss_gradient_dim2_mx <- matrix(0, ncol = B,
                                nrow = length(rho_finite_element_dim1$y))

for(i in 1:B){
  h_dim1 <- sample(x = c(-1,1), size = length(rho_finite_element_dim1$y), replace = TRUE)
  h_dim2 <- sample(x = c(-1,1), size = length(rho_finite_element_dim2$y), replace = TRUE)
  
  
  rho_fe_dim1_plus <- rho_finite_element_dim1
  rho_fe_dim2_plus <- rho_finite_element_dim2
  
  rho_fe_dim1_plus$y <- rho_finite_element_dim1$y + delta_dim1*h_dim1
  rho_fe_dim2_plus$y <- rho_finite_element_dim2$y + delta_dim2*h_dim2
  
  rho_fe_dim1_minus <- rho_finite_element_dim1
  rho_fe_dim2_minus <- rho_finite_element_dim2
  
  rho_fe_dim1_minus$y <- rho_finite_element_dim1$y - delta_dim1*h_dim1
  rho_fe_dim2_minus$y <- rho_finite_element_dim2$y - delta_dim2*h_dim2
  
  Loss_plus <- get_loss(rho_fe_dim1_plus,
                        rho_fe_dim2_plus,
                        rho_prev = rho_0)
  
  Loss_minus <- get_loss(rho_fe_dim1_minus,
                        rho_fe_dim2_minus,
                        rho_prev = rho_0)
  Loss_gradient_dim1 <- (Loss_plus - Loss_minus)/(2*delta_dim1*h_dim1)
  Loss_gradient_dim2 <- (Loss_plus - Loss_minus)/(2*delta_dim2*h_dim2)
  
  Loss_gradient_dim1_mx[,i] <- Loss_gradient_dim1
  Loss_gradient_dim2_mx[,i] <- Loss_gradient_dim2
  cat(paste("finish", i, "\n"))
}

```

```{r}
Loss_gradient_dim1_mean <- rowMeans(Loss_gradient_dim1_mx)

Loss_gradient_dim1_mean_sub <- ifelse(abs(Loss_gradient_dim1_mean) < 1e5, 
                                      Loss_gradient_dim1_mean,
                                      sign(Loss_gradient_dim1_mean)*1e5)
```

```{r}
hist(Loss_gradient_dim1_mean_sub)
```

```{r}
rho_fe_dim1_update <- rho_finite_element_dim1
rho_fe_dim1_update$y <- rho_finite_element_dim1$y - 1/(1e6)*Loss_gradient_dim1_mean_sub

plot(rho_fe_dim1_update)
```

```{r}
rho_fe_dim1_update <- rho_finite_element_dim1
rho_fe_dim1_update$y <- rho_finite_element_dim1$y - 1/(1e6)*Loss_gradient_dim1_mean

plot(y=abs(rho_fe_dim1_update$y[100:400]),x = rho_fe_dim1_update$x[100:400])
```
















